---
title: "Statistics Project - Spotify Data set"
author: "Johnny Solomovich"
date: '2022-06-01'
---
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
The project data set is named "Spotify 2010-1019 Top Hits list" and contains one thousand 
observations. It is taken from Kaggle.com. 
```{r, include=F}
library(tidyverse)
spotifyC <- read.csv("C:/Users/yoni5/Downloads/Spotify 2010  2019 Top 100.csv")
```
The main connection I will try to examine 
is between **songs genres** and **the amount of time it takes to become a hit**. 
The operational variables will be the main categorical genre's of songs and number 
of years it took for the song to enter the list. 

I created two new variables to examine my main connection. 
The first variable called "General Genre". It’s a categorical variable. Due to the fact 
that there are 89 different categorical options for the original genre in the data set, 
I congregate most of the observations by the four main categories (Rock, Pop, EDM and Hip-Hop)
and used Chosic.com categorization for it [9]. One hundred and seventy-four observations 
who did not fit to any of the general genres were omitted. 
```{r}
# Creating the General Genre test variable:
spotifyC$general_genre <- rep(NA,length(spotifyC$artist))

spotifyC$general_genre[which(grepl(".*(rock|indie|g funk|alternative metal).*",spotifyC$top.genre))] <- "rock"

spotifyC$general_genre[which(grepl(".*(pop|r&b).*",spotifyC$top.genre))] <- "pop"

spotifyC$general_genre[which(grepl(".*(hip hop|rap).*",spotifyC$top.genre))] <- "hip hop"

spotifyC$general_genre[which(grepl(".*(dance|electro|house|edm|big room).*",spotifyC$top.genre))] <- "dance"
```
One hundred and seventy-four observations 
who did not fit to any of the general genres were omitted:
```{r}
spotifyC <- na.omit(spotifyC)
```
The second variable called "Same.diff" and it’s a binary variable. 
It is a summation between the year the song was published and the year it entered the list 
songs who had the same year got 0 as their score. 
All other conditions got 1 as their score. Both years are part of the available data in the 
data set. 
```{r}
#Creating the Years To the Top variable:
spotifyC$years.to.top <- spotifyC$year.released - spotifyC$top.year
#Creating the same or different years test variable:
spotifyC$same.diff <-  as.numeric(cut(spotifyC$years.to.top,c(-Inf,0,Inf),c(0,1)))-1
```
Fifty-seven observations were omitted because of abnormal results (over 5 years or 
a negative result. Meaning the song had entered the list before it was published). 
Therefore, the new data set contains 763 observations.
```{r}
#Filtering the entire data from abnormal observations:
spotifyC <- na.omit(filter(spotifyC,!(spotifyC$years.to.top > 5 | spotifyC$years.to.top < 0)))
```
I'll try to estimate the different proportions of songs that entered the list
the same year they were released.
```{r}
pop     <- spotifyC[which(spotifyC$general_genre =="pop"),20]

rock    <- spotifyC[which(spotifyC$general_genre=="rock"),20]

hip_hop <- spotifyC[which(spotifyC$general_genre=="hip hop"),20]

dance   <- spotifyC[which(spotifyC$general_genre=="dance"),20]

```

```{r}
(results <- data.frame(Genre      = 
                        c("Pop","Rock","Hip Hop","Elcetro-Dance"),
                      N          = 
                        c(length((pop)),
                          length((rock)),
                          length((hip_hop)),
                          length((dance))),
                      Proportion = 
                        c(1-mean((pop)),
                        1-mean((rock)),
                        1-mean((hip_hop)), 
                        1-mean((dance))),
                      Same.year  = 
                        c(length((pop)[which(pop==0)]),
                         length((rock)[which(rock==0)]),
                         length((hip_hop)[which(hip_hop==0)]),
                         length((dance)[which(dance==0)])),
                      Diff.year  = 
                        c(length((pop)[which(pop>0)]),
                         length((rock)[which(rock>0)]),
                         length((hip_hop)[which(hip_hop>0)]),
                         length((dance)[which(dance>0)]))))
```
Results: We can see that the mode value is "dance", but it's also the lowest 
proportion of songs that entered the same year they got out. 
```{r}
dimnames(results)[[1]] <- c("Pop","Rock","Hip Hop", "Electro-Dance")
mosaicplot(results[,c(4,5)], 
           dir = c("v","h"), xlab = "Genre", ylab = "Proportions" ,
           main = "Spotify Genre top hits proportions",color=c("darkgoldenrod","darkcyan"))
```
Interpretation to the Graphs: Mosaic plot is used when we have 2 categorical variables. 
in our case it's songs general genres and same/different year to enter the list (binary).
We can see that the proportions between the number of songs in the same year 
(to enter the list) and the different year is different between the categories. 

We'll try to see if the difference is significant by the relevant test. 
We'll use Chi-Square test (2 categorical variables and with more than 2 categories
in at least one of them). 

```{r}
OBS = cbind(rbind(same.year = results[,4],diff.year = results[,5],Total = results[,2]),
  c(sum(results[,4]),sum(results[,5]),sum(results[,2])))
`colnames<-`(OBS,c("Pop","Rock","Hip Hop", "Electro-Dance","Total"))

EXP = matrix(NA,2,4)
  

for (i in 1:(length(OBS[,1])-1)){
  for (j in 1:(length(OBS[1,])-1)){
    EXP[i,j] = ((OBS[i,5]*OBS[3,j])/OBS[3,5])
  }
}
`colnames<-`(EXP,c("Pop","Rock","Hip Hop", "Electro-Dance"))

chisq.test(OBS[1:2,1:4],EXP,simulate.p.value = TRUE)
```
Test Result: We can see that P-value <= alpha. Therefore, we can reject the null 
hypothesis. We can say that songs genre and number of years to get to Spotify's 
Top Hits are dependent. 

Discussion: so far we know that there's a dependency between the genres and the amount of 
time to become a hit. Now I'll try to understand what exactly makes this difference 
between the songs genres. 

-------------------------

One research on Spotify's songs suggested that high danceability scores with low 
instrumentals scores related to songs success in songs charts [10]. 

We don't have a variable who is related directly to how much the song is focused on 
instrumentals, but we do have the dnce quantitative variable (description: How easy 
it is to dance to the song?). 

Additionally, from my intuition EDM and rock songs (the dance genre) using more
instrumentals than in pop and hip-hop songs.  

Therefore, we will try to see if there is a difference between songs general genres
by the dnce variable, while we try to see if the difference is according to my intuition
(rock and EDM will have different results than hip-hop and pop songs).

-------------------------

First of all, we need to see if the observations of dnce are distributed normally:
```{r}
#Normality Check:
ggplot(data.frame(y = spotifyC$dnce), aes(sample = y)) +
#  facet_wrap(~spotifyC$general_genre,nrow=2) +
  stat_qq() +  
  stat_qq_line() + 
    ggtitle("Normaility test for the dnce observations") +
  xlab("Z Quantities") + 
  ylab("Sampled Quantiles (dnce)") + 
  theme(plot.title=element_text(hjust=0.5))
```
Decision: The observations distributed different from the normal distribution. 
It's dispersed more than the normal distribution, so it could indicate the dnce 
observations distributed t with 686 df. I prefer not to do Chi-square for 
goodness of fit because it's not offers me much in return in this case. I'd rather
not assume anything on the distribution of the observation and do a-parametric test.

Therefore, I'll use Wilcoxon Rank Sum Test.

# Descriptive Statistics:
```{r}

```
# Graph:
```{r}
ggplot(spotifyC,aes(dnce,general_genre)) +
  geom_boxplot() + 
  ggtitle("dnce score by General Genre") +
  xlab("dnce score") + 
  ylab("Genre") + 
  coord_flip() +
  theme(plot.title=element_text(hjust=0.5))
```
Interpretation: We can see there is a difference between the different genres and
specifically between EDM and Rock  as opposed to hip-hop and pop songs. 

Now I'll try to see if there is a difference between EDM and 
pop or Hip-Hop and between Rock and Pop or Hip Hop. 

I'll also do Holm correction for each of the 4C2 combinations of the song's 
general genres. 
```{r}
P.value.dnce = c()

P.value.dnce[1] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"],spotifyC$dnce[spotifyC$general_genre=="rock"])$p.value

P.value.dnce[2] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"],spotifyC$dnce[spotifyC$general_genre=="dance"])$p.value

P.value.dnce[3] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"],spotifyC$dnce[spotifyC$general_genre=="hip hop"])$p.value

P.value.dnce[4] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="dance"],spotifyC$dnce[spotifyC$general_genre=="rock"])$p.value

P.value.dnce[5] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="dance"],spotifyC$dnce[spotifyC$general_genre=="hip hop"])$p.value

P.value.dnce[6] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="rock"],spotifyC$dnce[spotifyC$general_genre=="hip hop"])$p.value

data.frame(p.value = p.adjust(sort(P.value.dnce), method = "holm"),
           a = c(0.05/6,0.05/5,0.05/4,0.05/3,0.05/2,0.05/1), 
           reject = c(p.adjust(sort(P.value.dnce), method = "holm")
                      <=c(0.05/6,0.05/5,0.05/4,0.05/3,0.05/2,0.05/1)))
```
Results: We can see that the adjusted P-value <= adjusted alpha. Therefore, we can 
reject the null hypothesis for all, expect for (pop vs dance). 
We can say that the songs dnce expected values are different.

However, the order of the averages is not like the order of the proportions in the
results data frame. Therefore, I'll assume that there is another variable who is
influencing on the amount of time it takes to a song to become a hit. 

One of the qualitative variables in our data is pop (Popularity score of the song - 
not a ranking). Moreover, the average scores by general genres in at the same order
as the proportion from results (data frame). 
```{r,include=FALSE}
ggplot(spotifyC,aes(dnce,general_genre)) +
  facet_wrap(~spotifyC$same.diff,nrow = 2) +
  geom_boxplot() + 
  ggtitle("dnce score by General Genre") +
  xlab("dnce score") + 
  ylab("Genre") + 
  coord_flip() +
  theme(plot.title=element_text(hjust=0.5))


(dnce.avg = data.frame(Genre      = 
                        c("Pop","Rock","Hip Hop","Elcetro-Dance"),
                      N          = 
                        c(length((spotifyC[which(spotifyC$general_genre=="dance"),8])),
                          length((spotifyC[which(spotifyC$general_genre=="rock"),8])),
                          length((spotifyC[which(spotifyC$general_genre=="pop"),8])),
                          length((spotifyC[which(spotifyC$general_genre=="hip hop"),8]))),
                      dnce.avg = 
                        c(mean((spotifyC[which(spotifyC$general_genre=="dance"),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="rock"),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="pop"),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="hip hop"),8]))),
                      Same.year  = 
                        c(mean((spotifyC[which(spotifyC$general_genre=="dance" &
                                                       spotifyC$same.diff==0),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="rock" &
                                                       spotifyC$same.diff==0),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="pop" &
                                                       spotifyC$same.diff==0),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="hip hop" &
                                                       spotifyC$same.diff==0),8]))),
                      Diff.year  = 
                        c(mean((spotifyC[which(spotifyC$general_genre=="dance" &
                                                       spotifyC$same.diff==1),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="rock" &
                                                       spotifyC$same.diff==1),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="pop" &
                                                       spotifyC$same.diff==1),8])),
                          mean((spotifyC[which(spotifyC$general_genre=="hip hop" &
                                                       spotifyC$same.diff==1),8])))))
```
```{r,include=FALSE}
#Normality Tests:
ggplot(data.frame(y = spotifyC$dnce[which(spotifyC$same.diff==0)]), aes(sample = y)) +
  stat_qq() + 
  stat_qq_line() + 
    ggtitle("Normaility test for the dnce observations") +
  xlab("Z Quantities") + 
  ylab("Sampled Quantiles (dnce)") + 
  theme(plot.title=element_text(hjust=0.5))

ggplot(data.frame(y = spotifyC$dnce[which(spotifyC$same.diff==1)]), aes(sample = y)) +
  stat_qq() + 
  stat_qq_line() + 
    ggtitle("Normaility test for the dnce observations") +
  xlab("Z Quantities") + 
  ylab("Sampled Quantiles (dnce)") + 
  theme(plot.title=element_text(hjust=0.5))


P.value.dnce.0 = c()

P.value.dnce.0[1] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"
                                              &spotifyC$same.diff==0],
                                spotifyC$dnce[spotifyC$general_genre=="rock"&
                                                 spotifyC$same.diff==0])$p.value

P.value.dnce.0[2] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"
                                              &spotifyC$same.diff==0],
                                spotifyC$dnce[spotifyC$general_genre=="dance"
                                              &spotifyC$same.diff==0])$p.value

P.value.dnce.0[3] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"
                                              &spotifyC$same.diff==0],
                                spotifyC$dnce[spotifyC$general_genre=="hip hop"
                                              &spotifyC$same.diff==0])$p.value

P.value.dnce.0[4] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="dance"
                                              &spotifyC$same.diff==0],
                                spotifyC$dnce[spotifyC$general_genre=="rock"
                                              &spotifyC$same.diff==0])$p.value

P.value.dnce.0[5] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="dance"
                                              &spotifyC$same.diff==0],
                                spotifyC$dnce[spotifyC$general_genre=="hip hop"
                                              &spotifyC$same.diff==0])$p.value

P.value.dnce.0[6] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="rock"
                                              &spotifyC$same.diff==0],
                                spotifyC$dnce[spotifyC$general_genre=="hip hop"
                                              &spotifyC$same.diff==0])$p.value

data.frame(p.value = p.adjust(sort(P.value.dnce.0), method = "holm"),
           a = c(0.05/6,0.05/5,0.05/4,0.05/3,0.05/2,0.05/1), 
           reject = c(p.adjust(sort(P.value.dnce.0), method = "holm")
                      <=c(0.05/6,0.05/5,0.05/4,0.05/3,0.05/2,0.05/1)))


P.value.dnce.1 = c()

P.value.dnce.1[1] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"
                                              &spotifyC$same.diff==1],
                                spotifyC$dnce[spotifyC$general_genre=="rock"&
                                                 spotifyC$same.diff==1])$p.value

P.value.dnce.1[2] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"
                                              &spotifyC$same.diff==1],
                                spotifyC$dnce[spotifyC$general_genre=="dance"
                                              &spotifyC$same.diff==1])$p.value

P.value.dnce.1[3] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="pop"
                                              &spotifyC$same.diff==1],
                                spotifyC$dnce[spotifyC$general_genre=="hip hop"
                                              &spotifyC$same.diff==1])$p.value

P.value.dnce.1[4] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="dance"
                                              &spotifyC$same.diff==1],
                                spotifyC$dnce[spotifyC$general_genre=="rock"
                                              &spotifyC$same.diff==1])$p.value

P.value.dnce.1[5] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="dance"
                                              &spotifyC$same.diff==1],
                                spotifyC$dnce[spotifyC$general_genre=="hip hop"
                                              &spotifyC$same.diff==1])$p.value

P.value.dnce.1[6] = wilcox.test(spotifyC$dnce[spotifyC$general_genre=="rock"
                                              &spotifyC$same.diff==1],
                                spotifyC$dnce[spotifyC$general_genre=="hip hop"
                                              &spotifyC$same.diff==1])$p.value

data.frame(p.value = p.adjust(sort(P.value.dnce.1), method = "holm"),
           a = c(0.05/6,0.05/5,0.05/4,0.05/3,0.05/2,0.05/1), 
           reject = c(p.adjust(sort(P.value.dnce.1), method = "holm")
                      <=c(0.05/6,0.05/5,0.05/4,0.05/3,0.05/2,0.05/1)))
```
```{r}

 pop.avg = 
                        c(mean((spotifyC[which(spotifyC$general_genre=="dance"),15])),
                          mean((spotifyC[which(spotifyC$general_genre=="rock"),15])),
                          mean((spotifyC[which(spotifyC$general_genre=="pop"),15])),
                          mean((spotifyC[which(spotifyC$general_genre=="hip hop"),15])))
```
Therefore, I'll try to see if dnce 
scores are also correlated to pop scores. If there is a correlation, we could assume
that dnce scores are correlated with the popularity scores that Spotify gives to each
song. 
```{r}
#function that add the linear model equation to the plot:
lm_eqn <- function(df){
m <- lm(df[,1] ~ df[,2], df);
eq <- substitute("Least Squares:"~~italic(y) == a + b %.% italic(x),
list(a = format(unname(coef(m)[1]), digits = 2),
b = format(unname(coef(m)[2]), digits = 2)))
as.character(as.expression(eq))}
```
```{r}
#function that add Tukey model equation to the plot:
myTukeyRL <- function(x,y){
#Dividing the X vector to 3 quantiles:
quants <- quantile(x, c(1/3, 2/3), type = 6)
#Finding the Y-low and Y-high:
y_anchor <- c(median(y[x <= quants[1]]), median(y[x > quants[2]]))
#Finding the X-low and X-high:
x_anchor <- c(median(x[x <= quants[1]]), median(x[x > quants[2]]))
#Finding our slope(betaRL) and X-intercept(alphaRL):
betaRL <- (y_anchor[2] - y_anchor[1]) / (x_anchor[2] - x_anchor[1])
alphaRL <- median(y - betaRL * x)
#Returning the RL-variables:
return(c(betaRL, alphaRL))
}
```
Our Intercepts and the slopes for each line:
```{r}
lm(spotifyC$dnce~spotifyC$pop)
(lm_res <- myTukeyRL(spotifyC$dnce,spotifyC$pop))
```
```{r}
ggplot(spotifyC,aes(dnce,pop)) +
  geom_point(size = 2, alpha = 0.75) + 
  geom_abline(colour = "royalblue3", size = 0.75) +
  geom_abline(intercept=lm_res[2],slope=lm_res[1],
    colour = "firebrick3", size = 0.75) +
  geom_text(x = 60, y = 45,
          label = lm_eqn(data.frame(dnce = spotifyC$dnce,
                                    pop = spotifyC$pop)),
    colour = "royalblue3", parse = TRUE) +
  geom_text(x = 30, y = 68,
    label = as.character(as.expression(substitute("Resistant Line:"~~
    italic(y) == 69.83 + 0.083 %.% italic(x)))),
    colour = "firebrick3",parse = TRUE) +
  ggtitle("dnce score by pop scores") +
  xlab("dnce score") + 
  ylab("pop scores") + 
  theme(plot.title=element_text(hjust=0.5))
```

Now I'll do a residual charts to each line to see if they both suitable to explain 
this plot.
```{r, include=FALSE}
spotify.reg = data.frame(pop.short = (spotifyC[which(spotifyC$pop>=80),8]),
           dnce.short = (spotifyC[which(spotifyC$pop>=80),15]))

ggplot(spotify.reg,aes(dnce.short,pop.short)) +
  geom_point(size = 2, alpha = 0.75) + 
  geom_abline(colour = "royalblue3", size = 0.75) +
  geom_abline(intercept=lm_res[2],slope=lm_res[1],
    colour = "firebrick3", size = 0.75) +
  geom_text(x = 90, y = 88,
          label = lm_eqn(data.frame(dnce = spotifyC$dnce,
                                    pop = spotifyC$pop)),
    colour = "royalblue3", parse = TRUE) +
  geom_text(x = 90, y = 68,
    label = as.character(as.expression(substitute("Resistant Line:"~~
    italic(y) == 69.83 + 0.083 %.% italic(x)))),
    colour = "firebrick3",parse = TRUE) +
  ggtitle("dnce score by pop scores") +
  xlab("dnce score") + 
  ylab("pop scores") + 
  theme(plot.title=element_text(hjust=0.5))
```

```{r}
lm_dat <- data.frame(residuals=lm(spotifyC$dnce~spotifyC$pop)$residuals,
                     Yi=spotifyC$pop)

ggplot(lm_dat,aes(lm_dat[,2],lm_dat[,1])) +
geom_point(size = 2, alpha = 0.75) +
geom_hline(colour = "royalblue3", yintercept = 0) +
ggtitle("Residuals by pop scores") +
xlab("Residuals") +
ylab("pop scores") +
theme(plot.title=element_text(hjust=0.5))

RL.res <- spotifyC$pop - (lm_res[2] + lm_res[1] * spotifyC$dnce)

suppressWarnings(print(
ggplot(lm_dat ,aes(lm_dat[,2],RL.res)) +
geom_point(colour = "grey24", size = 2, alpha = 0.4) +
geom_hline(colour = "firebrick3", yintercept = 0) +
ggtitle("Resistant Line Residuals") +
ylab("Residuals") +
xlab("pop scores") +
theme(plot.title=element_text(hjust=0.5))
))
```
Interpretation: We can clearly see that there is a connection between the residuals
of the pop scores from the RL line, so we can't explain it with a linear model.
However, we can do it with the least squares line, even though it seems like it's 
not the best idea. Therefore, I'll at least try and test the line slope to see if 
it can explain a correlation between the dnce scores and the pop scores.
```{r}
summary(lm(spotifyC$dnce~spotifyC$pop))
```
Results: We got P-value < 0.01, so it's not that common to get this slope for this 
amount of observations.

However, it doesn't seem like it shows a real correlation between the two variables.
I'll try it by another way - I'll create Intervals of the dnce scores and see if I
can recognize connection between the variables. by the previous plot we can argue that
we see three different sub-groups of observations (by dnce scores) together: 40-60,
60-80 and 80-Inf. we'll see if there is a correlation of it with the proportions
from results (data-frame).

```{r}
results.2 <- data.frame(genre = c("dance","rock","pop","hip-hop"),
           n.40.60 = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>40 & spotifyC$pop<60),15])),
                       length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>40 & spotifyC$pop<60),15])),
                       length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>40 & spotifyC$pop<60),15])),
                      length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>40 & spotifyC$pop<60),15]))),
           n.60.80 = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=80),15]))),
           n.80.inf = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=80),15]))),
           p.60.Inf.vs.80.Inf = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=80),15]))/
                                   length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=60),15])),
                                 length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=80),15]))/
                                   length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=60),15])),
                                 length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=80),15]))/
                                   length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=60),15])),
                                 length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=80),15]))/
                                   length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=60),15]))
                                   ))
```
Now we can see that the ranking of the proportions from the first analysis is the same
as here. Most of the observations got 60 or more in the pop score, 
but The proportions in here suggest that the genres who were found with higher proportion
of songs that became a hit in their first year, has also more songs who got 80 or more 
in the pop scores and they are also a part of the interval [80-Inf) dnce scores.

I'll try to test those frequencies with the suitable test for it. For alpha of 0.05
I'll test if the pop scores and the dnce scores are dependant. the null-hypothesis 
is that they aren't.
```{r}
OBS.dnce.pop = data.frame(genre = c("dance","rock","pop","hip-hop"),
           n.60.80 = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=60 & spotifyC$pop<80),15]))),
           n.80.inf = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=80),15])),
                       length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=80),15]))),
           Total = c(length((spotifyC[which(spotifyC$general_genre=="dance"
                                      &spotifyC$pop>=60),15])),
           length((spotifyC[which(spotifyC$general_genre=="rock"
                                      &spotifyC$pop>=60),15])),
           length((spotifyC[which(spotifyC$general_genre=="pop"
                                      &spotifyC$pop>=60),15])),
           length((spotifyC[which(spotifyC$general_genre=="hip hop"
                                      &spotifyC$pop>=60),15]))))
```
```{r}
OBS.dnce.pop[5,] <- c(NA,length((spotifyC[which(spotifyC$pop>=60 & spotifyC$pop<80),15])),length((spotifyC[which(spotifyC$pop>=80),15])),
        length((spotifyC[which(spotifyC$pop>=60),15])))

EXP.dnce.pop = matrix(NA,5,3)
  

for (i in 1:(length(OBS.dnce.pop[,1])-1)){
  for (j in 2:(length(OBS.dnce.pop[1,])-1)){
    EXP.dnce.pop[i,j] = ((OBS.dnce.pop[i,4]*OBS.dnce.pop[5,j])/OBS.dnce.pop[5,4])
  }
}
`colnames<-`(OBS.dnce.pop,c("Pop","Rock","Hip Hop", "Electro-Dance"))

chisq.test(OBS.dnce.pop[1:4,2:3],EXP.dnce.pop[1:4,2:3],simulate.p.value = TRUE)

dimnames(results.2)[[1]] <- c("Electro-Dance","rock","Pop","Hip Hop")
mosaicplot(results.2[,c(3,4)], 
           dir = c("v","h"), xlab = "Genre", ylab = "Proportions" ,
           main = "Dependency between dnce scores and pop scores",
           color=c("darkgoldenrod","darkcyan"))
```
Test Result: We can see that P-value <= alpha. Therefore, we can reject the null 
hypothesis. We can say that songs pop scores and the dnce scores (by intervals) 
are dependent.

Summary: This projects is based on two parts. 
first of all, to show that the amount of time to become a hit song in Spotify depends on
the song genre. I calculated how many songs, from each general genre, were hits in 
their 1st year. In my surprise, EDM songs had the lowest proportion (I believed that based
on the fact it's a popular genre in this age, they'll had better score). Moreover, 
in my surprise pop songs weren't at the top either. So The second part was to try and 
find if there are variables in the data set who can show the same proportion of success 
from the first part (EDM, Rock, Pop, Hip-Hop) and then have some info to make more sensible
conclusions. I tried to show that songs general genre has different danceability (dnce) scores 
averages. The results showed it was right, but not in the way I suspected it would be 
(similar to the proportions before). I believed dnce scores do have an influence 
on those scores somehow, so I wanted to test it with another variable and chose the 
one who is suspicious enough to influence on the original proportions - popularity 
scores. In my opinion, even though the correlation slope had low p-value, I believed 
there's another option to see if there is a connection between the variables and the 
original proportions. I aggregated the observations by 3 intervals I created by 
the look of them in the plot and found that most of the observations has dnce score of 
60 and beyond. so I wanted to see how many of those are also in the 80 and beyond group.
Apparently, the proportions of the 80 and beyond from the rest are ranked the same as
the original proportions. 

In my opinion, it could suggest that from some dnce score threshold, you'll become
a hit song, but you also need a higher score in popularity to enter in the same 
year you got out. EDM songs has a lot of songs that will be considered danceable, 
but not all of them are big time hits who will enter the list in their 1st year.
It's noteworthy to point out that hip-hop songs has the best highest proportion
of success and the best proportion of most songs who has high dnce scores (80 or more) 
and pop scores (80 or more). It will be interesting to search why hip-hop songs 
are so popular and danceable to the point that most of the songs that got out also
became a hit in Spotify in their 1st year they where released. 
Additionally, It's important to comment that this all calculations of the 1st proportions
based on annual dates, so more detailed info could show us other results or even 
a dependent variable like periods of time of releasing songs in each genre that 
influenced on those results. Moreover, there is also the influence of the listeners 
age. music is consumed differently based on ages of the listeners [3], so it's reasonable
to believe it had influenced on the original proportions (maybe songs are popular because
younger people just using Spotify more?). I'm aware of the fact that my assumptions
here could be influenced by those potential variables, but in my opinion the overall 
assumption is still stand - Songs genres relates to songs popularity. But of course
the reasons to it are more intresting.


[1] https://www.chosic.com/list-of-music-genres/
[2] Al-Beitawi, Z., Salehan, M., & Zhang, S. (2020). What makes a song trend? Cluster analysis of musical attributes for Spotify top trending songs. Journal of Marketing Development and Competitiveness, 14(3), 79-91.
[3] https://www.statista.com/statistics/253915/favorite-music-genres-in-the-us/
